---
title: "Introduction to xdbi"
author: "Vehbi Sinan Tunalioglu"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Quickstart

Import the library:

```{r}
library(xdbi)
```

Establish a connection:

```{r, results="hide"}
connect(RSQLite::SQLite(), ":memory:")
```

Write a `data.frame` to a new table:

```{r, results="hide"}
data(mtcars)
writeTable("mtcars", mtcars)
```

Read an entire table from the database:

```{r}
myData <- readTable("mtcars")
head(myData)
```

Write the same data with row names:

```{r, results="hide"}
writeTable("mtcars2", mtcars, withRowNames=TRUE)
```

This will result in reading with row names:

```{r}
myData2 <- readTable("mtcars2")
head(myData2)
```

Now issue an SQL command and note that the `row_names` column is not
interpreted as real row names but just as an ordinary column:

```{r}
readData("SELECT * FROM mtcars2 WHERE cyl <= %d", 6)
```

We still have access to `DBI` interface:

```{r}
resultSet <- DBI::dbSendQuery(getConnection(), sprintf("SELECT * FROM mtcars2 WHERE cyl <= %d", 6))
DBI::dbFetch(resultSet, n=-1)
```

But, we need to clear the result set manually:

```{r, results="hide"}
DBI::dbClearResult(resultSet)
```

We can also import partial table as long as column names are a match
(now things are getting a bit more interesting finally):

```{r, results="hide"}
myInput <- data.frame(cyl=c(4,6,NA), disp=c(NA, NA, 300))
writeData("mtcars", myInput)
```

Then the resulting table would read:

```{r}
readTable("mtcars")
```

Disconnect the existing connection:

```{r, eval=FALSE, results="hide"}
disconnect()
```

## Working with rows

You can insert individual records. Let's start with a dummy table:

```{r}
myTable <- data.frame(no="12345", name="Alice", gpa=3.4)
```

Write the table to DB:

```{r, results="hide"}
writeTable("gpas", myTable)
```

Let's write a new record:

```{r, results="hide"}
insertRow("gpas", no=23456, name="Bob", gpa=1.2)
```

Is it there?

```{r}
readTable("gpas")
```

Yep... How about updating a row? The second argument in the below
statements are for finding out which rows to update.

```{r, results="hide"}
updateRow("gpas", list(no=12345), gpa=4.0)
updateRow("gpas", list(no=23456, name="Bob"), gpa=2.0, no="98765")
```

Changes?

```{r}
readTable("gpas")
```

You can `select` rows, too:

```{r}
selectRow("gpas", no=12345, .ordering="gpa desc")
```

We can check if any rows exists with the given where clause:

```{r}
existsRow("gpas", no=12345)
existsRow("gpas", no=00000)
```

Finally, we can do an `upsert`, ie. update if it exists:

```{r, results="hide"}
upsertRow("gpas", list(no=12345), gpa=1.0, name="Alice")
upsertRow("gpas", list(no=45678), gpa=4.0, name="Susan")
```

Hence:

```{r}
readTable("gpas")
```

```{r, echo=FALSE, results="hide"}
## End of vignette, close the connection:
disconnect()
```
